---
title: "Auscensus - Getting Data"
resource_files:
  - vignetttes/house-preference-flow*.png
author: "Carlos YANEZ SANTIBANEZ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auscensus - Getting Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning=  FALSE,
  collapse = TRUE, comment = "#>", out.width = "100%",
  fig.width = 7, fig.height = 4, dev = "CairoPNG", dpi = 150, fig.path = "vignettes/get-data-",
  message = FALSE, warning = FALSE, error = FALSE
)
library(ggplot2)
library(auscensus)

```

Once the data has been imported, the package is ready to pull data from the data packs.

## Figuring out what to retrieve

The first thing to figure out what data points are going to be retrieved. In order to answer that, the following questions need to be answered:

- Which statistics do I want to pull, i.e. which data point from which table?
- Which geographical structure do I want  (e.g. SAs, LGAs, electorates)? Do i want data for all of them or just a handful?
- For which years do I want the data?

To answer those questions, {auscensus} comes with a family of functions, all starting with *list_*.

To figure out where is the data point to retrieve, the first step to retrieve a list of available tables, which will return their code plus the year when they have been published. this can be achieved with *list_census_tables()* 

```{r}
list_census_tables()

```

This function comes with two in-built filters. The first one allows filtering by table number, whereas the second uses a regular expression to filter the table names

```{r}
list_census_tables("20")
```

```{r}
list_census_tables(table_name_regex = "[Cc]ountry")
```

Please note that even though the table names have changed from year to year, all tables with the same number contain the same data.

Once the table has been selected, *list_census_attributes()* can be used to select a data point. Since tables can contain many attributes, thus this function also allows filtering the results by a regular expression. For instance, if we pick table *04* (Age by Sex) and want to find the data points for 60-year-olds, plus the total, we can use the below command.

```{r}
list_census_attributes(number="04",attribute_regex = "60|[Tt]otal")
```

To get which geographical structures are available, *list_census_geo_types()* can help. Once selected an appropriate value, *list_census_geo()* can be used to search for particular divisions. In the example below, we are confirming if data is available for three inner-Melbourne Local Government areas.

```{r}
list_census_geo_types()
```

```{r}
list_census_geo("LGA",geo_name_regex = "[Mm]elbourne|[Ss]tonnington|Yarra")
```

Please note that the above functions only show data for the *available* data packs - if a file has not been imported, its associated year won't be presented in the results.

## Retrieving Census Data

Once the questions have been resolved, it is possible to use {auscensus} to retrieve the data with *get_census_data()*. This function will present the result for each {table, year, geography} combination as a different element within a list. By default the function does not return data but a parquet binding function - this can be changes by setting *collect_data = TRUE*. For example:

```{r}

data <- get_census_data(census_table = list_census_tables("04"),
                        geo_structure = "LGA",
                        collect_data = TRUE)
  
names(data)   

```

```{r}
data$`2006_LGA_B04`
```

Note: Be mindful some tables are quite large!

By default, this command will store each table in a parquet file in the cache by directory if you want to stop that from happening use the argument *cache=FALSE*. Similarly, its default behaviour is to read data from cached files before extracting them from the data pack. If you want to ignore the cache, use *ignore_cache=TRUE*.

## Getting ready to use summaries

Although *get_census_data()*, provides great flexibility for data processing, in most cases, it is just of interest to get just a table with a limited number of variables for analysis. For this purpose, this package provides *get_census_summary()*. For instance, if we are only interested in getting the numbers for 60-year-olds  by LGA in the 2021 census (see above), we can use:


```{r}

get_census_summary(table_number = "04",
                   attribute = list("60 year old male"=c("Age_years_60_males","Age (Years): 60_males"),
                                     "60 year old female"=c("Age_years_60_males","Age (Years): 60_females")),
                   geo_structure = "LGA")

```

Finally, this function contains two more parameters of interest: *geo_units* filters the results by giving a vector of names; *reference_total* passes the function a vector of the names of another attribute which will be used as a "total" value to calculate the attributes as a percentage of said total value. For instance, if we want to narrow the above results to the LGAs of Melbourne, Stonnington and Yarra, and present the results as a percentage of the total population, we have:

```{r}

get_census_summary(table_number = "04",
                   attribute = list("60 year old male"=c("Age_years_60_males","Age (Years): 60_males"),
                                     "60 year old female"=c("Age_years_60_males","Age (Years): 60_females")),
                   geo_unit_names  = c("Melbourne","Stonnington","Yarra"),
                   reference_total = list("Total"=c("Total_persons")),
                   geo_structure = "LGA")
```

## Creating attributes list

Sometimes lists of attributes can be very big, consolidating multiple statistics into one category (i.e. aggregate all places of birth by continent).
To facilitate the process, *attribute_tibble_to_list()* can take the two first columns of a data frame and convert them to the desired format:

```{r}
attributes <- tribble(~Census_stat, ~ Group,
                     "Age_years_60_males","60 year old male",
                     "Age (Years): 60_males","60 year old male",
                     "Age_years_60_males","60 year old female",
                    "Age (Years): 60_females","60 year old female")

attribute_tibble_to_list(attributes)

```
