---
title: "Auscensus - Setup"
author: "Carlos YANEZ SANTIBANEZ"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auscensus - Setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  warning=  FALSE,
  collapse = TRUE, comment = "#>", out.width = "100%",
  fig.width = 7, fig.height = 4, dev = "CairoPNG", dpi = 150, fig.path = "vignettes/setup-",
  message = FALSE, warning = FALSE, error = FALSE
)
library(auscensus)

```

{auscensus}'s goal is to provide an easy interface to access data from Australian Census data packs. However, each data pack is a large file and thus they have not been included in the package. "Metadata" has been uploaded [here](https://github.com/carlosyanez/auscensus/releases/tag/aux_files) and is retrieved by the package as needed but **data files are not included**. To get things ready for data extraction, follow the below steps.

## Download the data

The first step to get {auscensus} ready to work is to download the census data packs. Although the package contains a function to do so (*data_census_download()*), it is recommended to download the files manually (due to the large size). Data packs come in multiple formats - this package has been designed with specific versions, please download the versions shown below (you can retrieve the links using *census_datapacks()*). You can download only some of the files if you are interested only in a particular census.

```{r}
census_datapacks()
```

## Cache folder

Upon load, the package will create a folder where it will store all imported, downloaded and cached files. You can find its location by running *find_cache()* or *Sys.getenv("auscensus_cache_dir")*. If you want to use the same cache in different environments (i.e. when using [{renv}](https://rstudio.github.io/renv/index.html)), you can do it via *Sys.setenv()* or *usethis::edit_r_environ()*.

## Importing the data

Once downloaded the data files, you can import them into the cache folder by using *data_census_import()*  - just provide a vector with the full path of the data pack zip files.

## Managing the cache

As mentioned above, the cache will contain:

- Imported data packs
- Metadata parquet files, used to assist the data retrieval.
- Cache queries (for easy to use), in parquet format (see more in this [article](articles/getting_data)). 

To keep an eye on the size of the cache, you can use *data_census_info()*

If you want to delete files, you can use *data_census_delete()*. This command will accept a vector with path names (which you can get from *data_census_info()*). If no argument is provided, it will delete all files in the cache.
